\mytitle

### 2015-08-03 MSc_diss.tex
Updates in Neuman2012 review

There have been a number of studies of the latitudial and seasonal variations of alkyl nitrates in marine air (Blake et al., 2003), in the Arctic (Swanson et al., 2003) and in the
Antarctic (Beyersdorf et al., 2010)
\subsubsection{Observational studies}

@@

\subsubsection{Modelling studies}
A huge number of modelling studies has been done about alkyl nitrates. In particular, O3 and the RONO2 oxida- tion products formed from reactions between these NOx and VOC emissions are examined in detail...

\citep{Neuman2012}
Source characterization and photochemical processes in transported air masses have been analyzed by examining RONO2 formation from anthropogenic emissions of alkane precursors [Simpson et al., 2003; Madronich, 2006; Simpson et al., 2006; Reeves et al., 2007; Worton et al., 2010].
Reactive nitrogen species and volatile organic compounds (VOCs) were measured from an instrumented aircraft during daytime flights in the marine boundary layer downwind from the area of surfacing oil. A unique VOC mixture, where alkanes dominated the hydroxyl radical (OH) loss rate, was emitted into a clean marine environment, enabling a focused examination of O3 and RONO2 formation processes. In the atmospheric plume from DWH, the OH loss rate, an indicator of potential O3 formation, was large and dominated by alkanes with between 5 and 10 carbons per molecule (C5–C10). Observations showed that NOx was oxidized very rapidly with a 0.8 h lifetime, producing primarily C6–C10 RONO2 that accounted for 78\% of the reactive nitrogen enhancements in the atmospheric plume 2.5 h downwind from DWH. Both observations and calculations of RONO2 and O3 production rates show that alkane oxidation dominated O3 formation chemistry in the plume. Rapid and nearly complete oxidation of NOx to RONO2 effectively terminated O3 production, with O3 formation
yields of 6.0 plus minus 0.5 ppbv O3 per ppbv of NOx oxidized.
Eight RONO2 compounds with between one and five carbon atoms
per molecule were measured, and in a few samples the sum of C6–C8 RONO2.
NOx loss by OH + NO2 was negligible here, as demonstrated by the absence of HNO3 production in the plumes (Figure 4). Instead, the rapid NOx loss was caused by the reaction of peroxy radicals with NO.
Figure 4 illustrates temporal changes in reactive nitrogen partitioning in the DWH plume. In fresh plumes (e.g., Figure 4a), NOx accounted for over 80\% of plume NOy. In aged plumes (e.g., Figure 4b), NOx oxidation was nearly complete, but NOx, PAN, HNO3, and the sum of eight C1–C5 RONO2 accounted for only 22\% of plume NOy. This result differs dramatically from previous studies, where the sum of NOx + PAN + HNO3 accounted for over 96\% of NOy in urban and petrochemical plumes [Neuman et al., 2002; Ryerson et al., 2003] and for approximately 90\% of NOy in rural areas [Parrish et al., 1993]. In urban air masses, RONO2 often accounted for a few percent of NOy [O’Brien et al., 1997; Flocke et al.,1998; Schneider et al., 1998; Talbot et al., 2003; Ryerson et al., 2003], and up to 10–20\% of NOy in some studies [Day et al., 2003; Perring et al., 2010].
The largest contribution to the total calculated RONO2 production rate is from C6–C10 alkanes, which were responsible for 80\% of the RONO2 production.
OPE. In the DWH plume, OPE was 6.0 plus minus 0.5 ppbv/ppbv after 1 h of transport (Figure 3b), and similar to values measured in U.S. urban areas [Trainer et al., 1995; Neuman et al., 2009]. In contrast, petrochemical plumes with OH loss rates dominated by alkenes had OPEs ranging from 10 to 18 [Ryerson et al., 2003] (Figure 3b). On 10 June, O3 enhancements in the DWH plumes were less than 15 ppbv (Figure 5b), even after NOx was completely oxidized. Although the ratio of VOC to NOx was very large, O3 production was limited by OH oxidation of C5–C10 alkanes that preferentially formed RONO2 and by NOx levels less than 10 ppbv.

@@

\begin{figure}\label{fig:Neu2008}
\caption{1}
\includegraphics[width=.5\linewidth]{{D:/UEA/ENV-MB4X_Dissertation/ENV-MB4X_Dissertation_thesis/ENV-MB4X_Dissertation_thesis_latex/pics/Neu2008_Fig.2_MeONO2_OCT_colorbar}.png}
\end{figure}
\begin{figure}
\caption{1}
\includegraphics[width=.5\linewidth]{{D:/UEA/ENV-MB4X_Dissertation/ENV-MB4X_Dissertation_thesis/ENV-MB4X_Dissertation_thesis_latex/pics/Neu2008_Fig.2_MeONO2_OCT}.png}
\end{figure}
\begin{figure}
\caption{1}
\includegraphics[width=.5\linewidth]{{D:/UEA/ENV-MB4X_Dissertation/ENV-MB4X_Dissertation_thesis/ENV-MB4X_Dissertation_thesis_latex/pics/Neu2008_Fig.3_OCT_NOxdiff}.png}
\end{figure}
\begin{figure}
\caption{October monthly mean mixing ratios of (a) methyl nitrate ($MeONO_2$, $CH_3ONO_2$) and (c) $O_3$, and percent difference (perturbation - control) in (b) $NO_x$, (d) $O_3$. All averaged over the boundary layer (below 850 mb). Adapted from \citep{Neu2008}}
\includegraphics[width=.5\linewidth]{{D:/UEA/ENV-MB4X_Dissertation/ENV-MB4X_Dissertation_thesis/ENV-MB4X_Dissertation_thesis_latex/pics/Neu2008_Fig.3_OCT_O3contol_O3diff}.png}
\end{figure}

>>>>

There have been a number of studies of the latitudial and seasonal variations of alkyl nitrates in marine air (Blake et al., 2003), in the Arctic (Swanson et al., 2003) and in the
Antarctic (Beyersdorf et al., 2010)
\subsubsection{Observational studies}
Anthropogenic!!! Neuman
Source characterization and photochemical processes in transported air masses have been analyzed by examining RONO2 formation from anthropogenic emissions of alkane precursors [Simpson et al., 2003; Madronich, 2006; Simpson et al., 2006; Reeves et al., 2007; Worton
et al., 2010]

@@

\subsubsection{Modelling studies}
A huge number of modelling studies has been done about alkyl nitrates. In particular, O3 and the RONO2 oxidation products formed from reactions between these NOx and VOC emissions are examined in detail...

A study that enabled us to focus on examination of $O_3$ and $RONO_2$ formation processes has been conducted by \citep{Neuman2012}. After an explosion at the Deepwater Horizon (DWH) offshore oil platform on 20 April 2010 large amounts of oil from the DWH spill reached the ocean surface. Subsequent oil evaporation caused a release of little methane and other light alkanes to the atmosphere forming a $VOC$ mixture atypical for the remote marine environment. At the same the $NO_x$ levels were enhanced in the vicinity of the oil spill due to spill response efforts (ship exhaust, surface oil burning, flaring of recovered gas). Airborne measurements of reactive nitrogen species ($NO_y$ = $NO_x$+$HNO_3$+$PAN$(peroxyacetul nitrate at the bottom)+$RONO_2$) and $VOCs$ in the marine boundary layer during daytime downwind from the DWH area showed that oxidation of alkanes with between 5 and 10 carbons per molecule ($C_5$–$C_{10}$) dominated $O_3$ formation chemistry in the plume. This circumstance led to a rapid and nearly complete oxidation of $NO_x$ to $RONO_2$ with mostly 6-10 carbons per molecule ($C_6$–$C_{10}$) effectively terminating $O_3$ production. On 10 June, $O_3$ enhancements in the DWH plumes were less than 15 ppbv, even after $NO_x$ was completely oxidized.

$O_3$ production efficiencies in the DWH plume were 6 plus minus 0.5 ppbv $O_3$ per ppbv of $NO_x$ oxidized after 1 h of transport. These values are comparable with those measured in U.S. urban areas [Trainer et al., 1995; Neuman et al., 2009], but smaller than those measured in the petrochemical plumes with $OH$ loss rates dominated by alkenes (10-18 ppbv $O_3$ per ppbv of $NO_x$) [Ryerson et al., 2003].

Temporal changes in reactive nitrogen partitioning in the DWH plume (Figure \ref{fig:Neuman2012}) nicely illustrate that in fresh part of the plume (Figure ??? a) $NO_x$ accounted for over 80\% of plume $NO_y$, in aged plumes (Figure ??? b), $NO_x$ oxidation was nearly complete, but NOx, PAN, HNO3, and the sum of eight $C_5$–$C_{10}$ $RONO_2$ accounted for only 22\% of plume $NO_y$. It means that $C_6$–$C_{10}$ $RONO_2$ account for 78\% of the aged plume $NO_y$\citep{Neuman2012}. This result differs dramatically from previous studies, where the estimated contribution of $RONO_2$ to $NO_y$ was much lower: approximately 4\% in urban and petrochemical plumes [Neuman et al., 2002; Ryerson et al., 2003], approx. 10\% in rural areas [Parrish et al., 1993], a few percent in urban air masses [O’Brien et al., 1997; Flocke et al.,1998; Schneider et al., 1998; Talbot et al., 2003; Ryerson et al., 2003], and up to 10–20\% of NOy in some studies [Day et al., 2003; Perring et al., 2010].

\begin{figure}[h!]\label{fig:Neuman2012}
\includegraphics[width=.5\linewidth]{{D:/UEA/ENV-MB4X_Dissertation/ENV-MB4X_Dissertation_thesis/ENV-MB4X_Dissertation_thesis_latex/pics/Neuman2012_Fig.4}.png}
\caption{Time series of reactive nitrogen species measured on 10 June 2010 in the MBL in plumes (a) 10 km and (b) 50 km downwind from DWH. $NO_x$, $NO_y$, and $HNO_3$ were measured once per s and $PAN$ was measured once per 2 s. $C_1$–$C_5$ alkyl nitrates were measured in whole air samples at locations indicated by the circles. The sum of alkyl nitrates (black lines) is inferred from the difference between $NO_y$ and ($NO_x$+$HNO_3$+$PAN$) \citep{Neuman2012}.}
\end{figure}

@@

\begin{figure}\label{fig:Neu2008}
\caption{1}
\includegraphics[width=.5\linewidth]{{D:/UEA/ENV-MB4X_Dissertation/ENV-MB4X_Dissertation_thesis/ENV-MB4X_Dissertation_thesis_latex/pics/Neu2008_Fig.2_MeONO2_OCT_colorbar}.png}
\end{figure}
\begin{figure}
\caption{1}
\includegraphics[width=.5\linewidth]{{D:/UEA/ENV-MB4X_Dissertation/ENV-MB4X_Dissertation_thesis/ENV-MB4X_Dissertation_thesis_latex/pics/Neu2008_Fig.2_MeONO2_OCT}.png}
\end{figure}
\begin{figure}
\caption{1}
\includegraphics[width=.5\linewidth]{{D:/UEA/ENV-MB4X_Dissertation/ENV-MB4X_Dissertation_thesis/ENV-MB4X_Dissertation_thesis_latex/pics/Neu2008_Fig.3_OCT_NOxdiff}.png}
\end{figure}
\begin{figure}
\includegraphics[width=.5\linewidth]{{D:/UEA/ENV-MB4X_Dissertation/ENV-MB4X_Dissertation_thesis/ENV-MB4X_Dissertation_thesis_latex/pics/Neu2008_Fig.3_OCT_O3contol_O3diff}.png}
\caption{October monthly mean mixing ratios of (a) methyl nitrate ($MeONO_2$, $CH_3ONO_2$) and (c) $O_3$, and percent difference (perturbation - control) in (b) $NO_x$, (d) $O_3$. All averaged over the boundary layer (below 850 mb). Adapted from \citep{Neu2008}}
\end{figure}
